Redundant calibration as a complex optimization problem

### 2017-05-05 paper.tex
implemented most of the fixes Gianni, suggested

\begin{abstract}
%In this paper we show that instead of splitting our redundant calibration problem into its phase-and-amplitude components (the more traditional approach) we can treat the complex unknown model parameters and their complex conjugates as separate unknowns. We also derive the Levenberg-Maquardt and Gauss-Newton update steps associated with this alternative parameter-and-conjugate framework. We focus more on the Levenberg-Maquardt algorithm in this paper as it exhibits better convergence properties when compared to Gauss-Newton. The Gauss-Newton algorithm is currently the de facto standard that is being used to perform redundant calibration. It is common practice to implement redundant calibration via a hierarchical algorithmic strategy. Using the Levenberg-Maquardt algorithm, therefore, reduces the number of algorithmic layers that are needed to perform redundant calibration. We also compare two previously proposed methods to accelerate redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method. We found that the alternating direction implicit method outperforms the preconditioned conjugate gradient method. However, the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases. We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method and the Levenberg-Marquardt algorithm with one another.
Observations of the redshifted 21-cm line from the epoch of reionization have recently motivated the construction of low frequency radio arrays with highly redundant configurations that provide an alternative calibration strategy - ''redundant calibration" - and boosts sensitivity on specific scales. We show that redundant calibration can be formulated in the domain of complex functions of a complex variable. This framework reduces the

has driven advancements in the calibration of radio interferometers and 
Calibration of the new generation of low frequency radio interferometers faces new challenges due to their very wide field of views 
The construction of new radio interferometers, particularly operating at low frequencies and 
In this paper we show that instead of splitting our redundant calibration problem into its phase-and-amplitude components (the more traditional approach) we can treat the complex unknown model parameters and their complex conjugates as separate unknowns. We also derive the Levenberg-Maquardt and Gauss-Newton update steps associated with this alternative parameter-and-conjugate framework. We focus more on the Levenberg-Maquardt algorithm in this paper as it exhibits better convergence properties when compared to Gauss-Newton. The Gauss-Newton algorithm is currently the de facto standard that is being used to perform redundant calibration. It is common practice to implement redundant calibration via a hierarchical algorithmic strategy. Using the Levenberg-Maquardt algorithm, therefore, reduces the number of algorithmic layers that are needed to perform redundant calibration. We also compare two previously proposed methods to accelerate redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method. We found that the alternating direction implicit method outperforms the preconditioned conjugate gradient method. However, the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases. We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method and the Levenberg-Marquardt algorithm with one another.
\end{abstract}

@@

We start the paper by presenting some useful preliminaries in Section~\ref{sec:preliminaries}. In Section~\ref{sec:p_statement}, we formulate redundant calibration 
as a least squares problem, i.e. Eq~\eqref{eq:least_squares_red}. In Section~\ref{sec:main_sec} we use the complex framework proposed by \citet{Smirnov2015} to derive the GN and LM
update rules (see Eq.~\eqref{eq:GN_update} to Eq.~\eqref{eq:update}), which we can use to solve Eq~\eqref{eq:least_squares_red}. In Section~\ref{sec:results}, we compare the computational complexity of two recently proposed redundant calibration speedup algorithms with one another.
We end the paper with some final conclusions.

@@

\subsection{Real and imaginary approach}
\label{sec:ri}
Using least-squares to solve complex problems can be problematic as $\bv$ needs to be analytic in its argument. In the case 
of Eq.~\eqref{eq:least_squares_red}, \citet{Liu2010} circumvents this difficulty by 
splitting the problem into its phase-and-amplitude components.

@@

\subsection{Complex Approach}
\label{sec:w}

@@

Analogous to \cite{Smirnov2015}, we recast Eq.~\eqref{eq:least_squares_red} to 
\begin{equation}
\label{eq:least_squares_complex}
\min_{\breve{\bz}} \|\breve{\br}\| = \min_{\breve{\bz}} \|\breve{\bd} - \breve{\bv}(\breve{\bz})\|, 
\end{equation}
where $\breve{\bz} = [\bz^T,\conj{\bz}^T]^T$. Eq.~\eqref{eq:least_squares_complex} is the complex conjugate augmented counterpart of Eq.~\eqref{eq:least_squares_red}.
The fact that $\bv$ is now analytic in its argument has the immediate consequence: it becomes possible to apply complex GN
and complex LM to Eq.~\eqref{eq:least_squares_complex}.

\subsubsection{Complex GN and LM}
\label{sec:complex_GN_LM}
The complex Jacobian associated with Eq.~\eqref{eq:least_squares_complex} is equal to
\begin{equation}
\label{eq:Jacobian}
\bJ = \begin{bmatrix}
       \bmJ & \bmJ^*\\
       \conj{\bmJ}^* & \conj{\bmJ} 
      \end{bmatrix},
\end{equation}
where 
\begin{equation}
[\bmJ]_{\alpha_{pq},i} = \begin{cases} 
     \frac{\partial v_{pq}}{\partial g_i} & \textrm{if}~i\leq N \\
     \frac{\partial v_{pq}}{\partial y_{i-N}} & \textrm{otherwise}  
\end{cases}, %, & \bmJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
\end{equation}
and
\begin{equation}
[\bmJ^*]_{\alpha_{pq},i} = \begin{cases} 
     \frac{\partial v_{pq}}{\partial \conj{g}_i} & \textrm{if}~i\leq N \\
     \frac{\partial v_{pq}}{\partial \conj{y}_{i-N}} & \textrm{otherwise}  
\end{cases}. %, & \bmJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
\end{equation}
The GN update step is now defined as:
\begin{equation}
\label{eq:GN_update}
 \Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br}.
\end{equation}
The LM update rule is defined in a very similar way:
\begin{equation}
\label{eq:LM_update}
\Delta \breve{\bz} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
\end{equation}

As mentioned in Section We use Eq.~\eqref{eq:GN_update} or Eq.~\eqref{eq:LM_update} to iteratively update our 
parameter vector as follows:
\begin{equation}
\label{eq:update}
\breve{\bz}_{k+1} = \breve{\bz}_{k} + \Delta \breve{\bz}_{k}. 
\end{equation}
We update our parameter vector by using Eq.~\eqref{eq:update} until convergence is reached. To improve readability we do not explicitly denote the iteration subscript $k$ in 
Eq.~\eqref{eq:GN_update} and Eq.~\eqref{eq:LM_update}.

@@

\section{Comparing ADI and PCG}
\label{sec:results}
In Section~\ref{sec:main_sec} we discussed the complex GN and LM algorithms and how they relate to redundant calibration. If we inspect Eq.~\eqref{eq:GN_update} and 
Eq.~\eqref{eq:LM_update} we see that the most expensive operation used by the GN and LM algorithms is the inversion of the Hessian, $\bH$. The most logical 
way of speeding up the algorithm is to investigate whether we can do this inversion in some efficient way.

@@

\subsubsection{Signal-to-noise Ratio}
\label{sec:snr}
The following SNR definition is used in this paper  
\begin{equation}
\label{}
\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
\end{equation}
where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.

@@

Since the diagonal entries of the $\bH$, associated with redundant calibration, are more significant than its off-diagonal entries it makes 
sense to try and approximate it with its diagonal (see Fig.~\ref{fig:hessian} and Eq.~\eqref{eq:red_H}).
If we substitute $\bJ^H\bJ$ with $\widetilde{\bH}$ and replace $\breve{\br}$ with $\breve{\bd} - \breve{\bv}$ in Eq.~\eqref{eq:GN_update} we obtain
\begin{equation}
\label{eq:appr_GN}
 \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H(\breve{\bd}-\breve{\bv})
\end{equation}
Utilizing the second identity in Eq.~\eqref{eq:identities} allows us to simplify Eq.~\eqref{eq:appr_GN} to
\begin{equation}
  \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}-\breve{\bz}.
\end{equation}
It now trivially follows that
\begin{equation}
 \breve{\bz}_{k+1} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}.
\end{equation}
We obtain a similar result if we repeat the above procedure, only in this case we use Eq.~\eqref{eq:LM_update} instead. Thus,
\begin{align}
\breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
 &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k. \label{eq:alpha}  
\end{align}
The analytic expression of $\bJ^H\breve{\bd}$ will be very similar to the analytic 
expression of $\bJ^H\breve{\br}$, the only difference being that in Eq.~\eqref{eq:ab} the letter $r$ would be replaced by a $d$. If we substitute the analytic expression
of $\bJ^H\breve{\bd}$ and $\widetilde{\bH}^{-1}$ (which is easily computed as the matrix is diagonal) into Eq.~\eqref{eq:alpha} we obtain the following two update rules:
\begin{equation}
\label{eq:g_update}
g_{i}^{k+1} = \alpha \frac{\sum_{j\neq i} g_j^k \widetilde{y}_{ij}^{~\!\!k} d_{ij}}{\sum_{j\neq i} |g_j^k|^2|y_{\zeta_{ij}}^k|^2} + (1-\alpha) g_i^k, 
\end{equation}
and
\begin{equation}
\label{eq:y_update}
y_{i}^{k+1} = \alpha \frac{\sum_{pq \in \mathcal{PQ}_i} \conj{g}_p^k g_q^k d_{pq}}{\sum_{pq \in \mathcal{PQ}_i}|g_p^k|^2|g_q^k|^2} + (1-\alpha) y_i^k. 
\end{equation}
The computational complexity of inverting $\widetilde{\bH}$ is $\mathbb{O}(P)$.

@@

\subsection{Preconditioned Conjugate Gradient Method}
\label{sec:pcg}
In this section we investigate the computational complexity of the PCG method when applied to the redundant calibration problem. 
As we mentioned in Section~\ref{sec:conj_grad} there are two factors which influence the computational complexity of the PCG method, namely the spectral condition number and the sparsity of the matrix we wish to invert (see Eq.~\eqref{eq:cg_bound}). 
We used a $\lambda$ value of two, implying we used Eq.~\eqref{eq:LM_update} instead of Eq.~\eqref{eq:GN_update} throughout this section.
We discuss the spectral condition number and the sparsity of the modified Hessian $\bmH$, which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity} respectively.
It is interesting to note, that the Hessian $\bH$ itself is singular which means that if we are to use Eq.~\eqref{eq:GN_update} instead we would need to be able to invert a singular matrix. In Section~\ref{sec:normal},
we showed that the PCG method would in fact be able to handle such a scenario.

\subsubsection{Spectral Condition Number}
\label{sec:scn}
In Section~\ref{sec:precon} we mentioned that the convergence properties of CG can be dramatically improved through preconditioning. Inspecting Fig.~\ref{fig:hessian} reveals that the Jacobian preconditioner
is a good candidate for the redundant use case. In Fig.~\ref{fig:kappa} we see that this choice of preconditioner performs very well, the value of the spectral condition number $\kappa$
becomes constant and independent of problem size, i.e. the number of antennas in the array. Preconditioning therefore effectively eliminates $\kappa$ from Eq.~\eqref{eq:cg_bound}. This result
is confirmed by Fig.~\ref{fig:cg_itr}. Fig.~\ref{fig:cg_itr} shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bmH$ is independent of the number of antennas in the array.
Moreover, the number of major iterations required is also much smaller than the dimension of $\bmH$.

@@

We can now use the sparsity factor to compute the order of the computational complexity of inverting $\bmH$ (assuming the effect of $\kappa$ is negliable) with
\begin{equation}
P^{c} = (1 - \gamma)P^2.
\end{equation}
Solving for $c$ we obtain
\begin{align}
c &= \log_{P}(1 - \gamma) + 2 & c_{\infty} &= \lim_{N\rightarrow \infty} c = 2. \label{eq:c}
\end{align}

@@

\subsection{Conjugate Gradient Method}
\label{sec:conj_grad}
The CG method was independently discovered by \citet{Hestenes1973} and \citet{Stiefel1952}. They later published a joint paper, which is now considered as the seminal
reference on CG \citep{Hestenes1952}. The iterative CG method is used to solve a particular class of linear system. The CG method is generally used to solve
\begin{equation}
\label{eq:linear_system}
\bb = \bA\bx,
\end{equation}
where $\bb\in\mathbb{C}^P$ is a known complex column vector of size $P$, $\bx\in\mathbb{C}^P$ is an unknown complex column vector with the same length as $\bb$, and $\bA\in\mathbb{C}^{P\times P}$ is a square positive-definite Hermitian matrix with the appropriate dimensions.  
Moreover, using the CG method to solve Eq.~\eqref{eq:linear_system} will only be computationally advantageous if $\bA$ is also sparse (contain many zero entries).
Using the iterative CG method to solve large sparse linear sytems was popularized by \citet{Reid1971}. A good tutorial on the CG method can be found in \citep{Shewchuk1994}.
The iterative CG method is presented in Algorithm \ref{algo:CG}.

@@

\subsubsection{Computational Complexity}
\citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
\citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of a major outer loop (line 6 of Algorithm \ref{algo:CG}). It turns out that at worst 
the number of major iterations that CG requires to converge\footnote{If the $\bA$-norm is used to calculate the error.} is proportional to $\sqrt{\kappa}$, where $\kappa$ is the spectral condition number of the matrix $\bA$.
The mathematical definition of $\kappa(\bA)$ is
\begin{equation}
\label{eq:kappa}
\kappa(\bA) = \frac{\lambda_{\textrm{max}}}{\lambda_{\textrm{min}}}, 
\end{equation}
where $\lambda_{\textrm{max}}$ and $\lambda_{\textrm{min}}$ respectively denote the largest and the smallest eigenvalue of $\bA$.
Moreover, the most expensive operation that is conducted within each major loop of Algorithm~\ref{algo:CG} is the vector-matrix product $\bA\bp$, which has a complexity of $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in $\bA$.
The computational complexity of CG is therefore 
\begin{equation}
\label{eq:cg_bound}
\mathbb{O}(\sqrt{\kappa}m). 
\end{equation}
It should now also be clear to the reader why the CG method is especially applicable to sparse linear systems. In sparse linear systems $m < P^2$.

\subsubsection{Preconditioning}
\label{sec:precon}
Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
We can now indirectly solve for $\bx$ by using 
\begin{equation}
\label{eq:prec}
\bM^{-1}\bA\bx = \bM^{-1}\bb.
\end{equation}
If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
of CG (see Eq.~\eqref{eq:cg_bound}). In this paper, we denote matrix inversion with $(*)^{-1}$. The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the identity matrix are equal to one).
If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$. However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place. 
In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).

>>>>

\begin{abstract}
%In this paper we show that instead of splitting our redundant calibration problem into its phase-and-amplitude components (the more traditional approach) we can treat the complex unknown model parameters and their complex conjugates as separate unknowns. We also derive the Levenberg-Maquardt and Gauss-Newton update steps associated with this alternative parameter-and-conjugate framework. We focus more on the Levenberg-Maquardt algorithm in this paper as it exhibits better convergence properties when compared to Gauss-Newton. The Gauss-Newton algorithm is currently the de facto standard that is being used to perform redundant calibration. It is common practice to implement redundant calibration via a hierarchical algorithmic strategy. Using the Levenberg-Maquardt algorithm, therefore, reduces the number of algorithmic layers that are needed to perform redundant calibration. We also compare two previously proposed methods to accelerate redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method. We found that the alternating direction implicit method outperforms the preconditioned conjugate gradient method. However, the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases. We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method and the Levenberg-Marquardt algorithm with one another.
Observations of the redshifted 21-cm line from the epoch of reionization have recently motivated the construction of low frequency radio arrays with highly redundant configurations that provide an alternative calibration strategy - ''redundant calibration" - and boosts sensitivity on specific scales. We show that redundant calibration can be formulated in the domain of complex functions of a complex variable. This framework reduces the 
\textbf{[GIANNI TO REWRITE AND FIX ABSTRACT]}
has driven advancements in the calibration of radio interferometers and 
Calibration of the new generation of low frequency radio interferometers faces new challenges due to their very wide field of views 
The construction of new radio interferometers, particularly operating at low frequencies and 
In this paper we show that instead of splitting our redundant calibration problem into its phase-and-amplitude components (the more traditional approach) we can treat the complex unknown model parameters and their complex conjugates as separate unknowns. We also derive the Levenberg-Maquardt and Gauss-Newton update steps associated with this alternative parameter-and-conjugate framework. We focus more on the Levenberg-Maquardt algorithm in this paper as it exhibits better convergence properties when compared to Gauss-Newton. The Gauss-Newton algorithm is currently the de facto standard that is being used to perform redundant calibration. It is common practice to implement redundant calibration via a hierarchical algorithmic strategy. Using the Levenberg-Maquardt algorithm, therefore, reduces the number of algorithmic layers that are needed to perform redundant calibration. We also compare two previously proposed methods to accelerate redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method. We found that the alternating direction implicit method outperforms the preconditioned conjugate gradient method. However, the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases. We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method and the Levenberg-Marquardt algorithm with one another.
\textbf{[Could we add JS Kenyon and SJ Wijnholds to author list?]}
\end{abstract}

@@

In Section~\ref{sec:sky_wirtinger}, we review the skymodel-based complex-and-conjugate calibration approach proposed by \citet{Smirnov2015}.
In Section~\ref{sec:red_wirtinger}, we show that this same approach can also be used to perform redundant calibration.
In Section~\ref{sec:results}, we compare the computational complexity of two recently proposed redundant calibration speedup algorithms with one another.
We end the paper with some final conclusions.

@@

\subsection{Real and Imaginary Approach}
\label{sec:ri}
Using least-squares to solve complex problems can be problematic as $\bv$ needs to be analytic in its argument. In the case 
of Eq.~\eqref{eq:least_squares_red}, \citet{Liu2010} circumvents this difficulty by 
splitting the problem into its phase-and-amplitude components.

@@

\subsection{Wirtinger Approach}
\label{sec:w}

@@

Analogous to \cite{Smirnov2015}, we recast Eq.~\eqref{eq:least_squares_red} to 
\begin{equation}
\label{eq:least_squares_complex}
\min_{\breve{\bz}} \|\breve{\br}\| = \min_{\breve{\bz}} \|\breve{\bd} - \breve{\bv}(\breve{\bz})\|, 
\end{equation}
where $\breve{\bz} = [\bz^T,\conj{\bz}^T]^T$. Eq.~\eqref{eq:least_squares_complex} is the complex conjugate augmented counterpart of Eq.~\eqref{eq:least_squares_red}.
%The fact that $\bv$ is now analytic in its argument has the immediate consequence: it becomes possible to apply complex GN
%and complex LM to Eq.~\eqref{eq:least_squares_complex}.

\subsubsection{Complex GN and LM}
\label{sec:complex_GN_LM}
The complex Jacobian associated with Eq.~\eqref{eq:least_squares_complex} is equal to
\begin{equation}
\label{eq:Jacobian}
\bJ = \begin{bmatrix}
       \bmJ & \bmJ^*\\
       \conj{\bmJ}^* & \conj{\bmJ} 
      \end{bmatrix},
\end{equation}
where 
\begin{equation}
[\bmJ]_{\alpha_{pq},i} = \begin{cases} 
     \frac{\partial v_{pq}}{\partial g_i} & \textrm{if}~i\leq N \\
     \frac{\partial v_{pq}}{\partial y_{i-N}} & \textrm{otherwise}  
\end{cases}, %, & \bmJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
\end{equation}
and
\begin{equation}
[\bmJ^*]_{\alpha_{pq},i} = \begin{cases} 
     \frac{\partial v_{pq}}{\partial \conj{g}_i} & \textrm{if}~i\leq N \\
     \frac{\partial v_{pq}}{\partial \conj{y}_{i-N}} & \textrm{otherwise}  
\end{cases}. %, & \bmJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
\end{equation}
The GN update step is now defined as:
\begin{equation}
\label{eq:GN_update}
 \Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br}.
\end{equation}
The LM update rule is defined in a very similar way:
\begin{equation}
\label{eq:LM_update}
\Delta \breve{\bz} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
\end{equation}
As was the case in Section~\ref{sec:sky_wirtinger}, we can now use Eq.~\eqref{eq:GN_update} or Eq.~\eqref{eq:LM_update} to iteratively update our 
parameter vector as follows:
\begin{equation}
\label{eq:update}
\breve{\bz}_{k+1} = \breve{\bz}_{k} + \Delta \breve{\bz}_{k}. 
\end{equation}
The analytic expressions for $\bJ$, $\bH$ and $\bJ^H\breve{\br}$ can be found in Appendix~\ref{sec:analytic}.

@@

\section{Comparing ADI and PCG}
\label{sec:results}
In Section~\ref{sec:red_wirtinger} we discussed the complex GN and LM algorithms and how they relate to redundant calibration. If we inspect Eq.~\eqref{eq:GN_update} and 
Eq.~\eqref{eq:LM_update} we see that the most expensive operation used by the GN and LM algorithms is the inversion of the Hessian, $\bH$. The most logical 
way of speeding up the algorithm is to investigate whether we can do this inversion in some efficient way.

@@

Furthermore, the following SNR definition is used in this paper  
\begin{equation}
\label{}
\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
\end{equation}
where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.

@@

Since the diagonal entries of the $\bH$, associated with redundant calibration, are more significant than its off-diagonal entries it makes 
sense to try and approximate it with its diagonal (see Fig.~\ref{fig:hessian} and Eq.~\eqref{eq:red_H}).
If we substitute $\bJ^H\bJ$ with $\widetilde{\bH}$ and replace $\breve{\br}$ with $\breve{\bd} - \breve{\bv}$ in Eq.~\eqref{eq:GN_update} we obtain
\begin{equation}
\label{eq:appr_GN}
 \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H(\breve{\bd}-\breve{\bv})
\end{equation}
Utilizing the second identity in Eq.~\eqref{eq:identities} allows us to simplify Eq.~\eqref{eq:appr_GN} to
\begin{equation}
  \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}-\breve{\bz}.
\end{equation}
It now trivially follows that
\begin{equation}
 \breve{\bz}_{k+1} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}.
\end{equation}
We obtain a similar result if we repeat the above procedure, only in this case we use Eq.~\eqref{eq:LM_update} instead. Thus,
\begin{align}
\breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
 &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k. \label{eq:alpha}  
\end{align}
The analytic expression of $\bJ^H\breve{\bd}$ will be very similar to the analytic 
expression of $\bJ^H\breve{\br}$, the only difference being that in Eq.~\eqref{eq:ab} the letter $r$ would be replaced by a $d$. If we substitute the analytic expression
of $\bJ^H\breve{\bd}$ and $\widetilde{\bH}^{-1}$ (which can easily be constructed using Appendix~\ref{sec:analytic}) into Eq.~\eqref{eq:alpha} we obtain the following two update rules:
\begin{equation}
\label{eq:g_update}
g_{i}^{k+1} = \alpha \frac{\sum_{j\neq i} g_j^k \widetilde{y}_{ij}^{~\!\!k} d_{ij}}{\sum_{j\neq i} |g_j^k|^2|y_{\zeta_{ij}}^k|^2} + (1-\alpha) g_i^k, 
\end{equation}
and
\begin{equation}
\label{eq:y_update}
y_{i}^{k+1} = \alpha \frac{\sum_{pq \in \mathcal{PQ}_i} \conj{g}_p^k g_q^k d_{pq}}{\sum_{pq \in \mathcal{PQ}_i}|g_p^k|^2|g_q^k|^2} + (1-\alpha) y_i^k. 
\end{equation}
The computational complexity of inverting $\widetilde{\bH}$ is $\mathbb{O}(P)$.

@@

\subsection{Preconditioned Conjugate Gradient Method}
\label{sec:pcg}
In this section we investigate the computational complexity of the PCG method when applied to the redundant calibration problem. 
We briefly review the PCG method in Appendix~\ref{sec:conj_grad}. If the reader is not that familiar with the PCG method we encourage 
the reader to briefly review Appendix~\ref{sec:conj_grad} before continuing with the paper. The PCG method can be used to efficiently invert Hermitian sparse 
matrices, which is exactly the kind of matrix $\bH$ is (see Fig.~\ref{fig:hessian} and Eq.~\ref{eq:red_H}). As we mention in Appendix~\ref{sec:conj_grad} there are two factors which influence the computational complexity of the PCG method, namely the spectral condition number $\kappa$ and the sparsity of the matrix we wish to invert (see Eq.~\eqref{eq:cg_bound}).

We used a $\lambda$ value of two, implying we used Eq.~\eqref{eq:LM_update} instead of Eq.~\eqref{eq:GN_update} throughout this section.
We investigate the spectral condition number and the sparsity of the modified Hessian $\bmH$, which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity} respectively.
It is interesting to note, that the Hessian $\bH$ itself is singular which means that if we are to use Eq.~\eqref{eq:GN_update} instead we would need to be able to invert a singular matrix. In Appendix~\ref{sec:normal},
we mention that the PCG method would in fact also be able to pseudo-invert $\bH$ directly.

\subsubsection{Spectral Condition Number}
\label{sec:scn}
In Appendix~\ref{sec:precon} we mentioned that the convergence properties of CG can be dramatically improved through preconditioning. Inspecting Fig.~\ref{fig:hessian}, reveals that the Jacobian preconditioner
is a good candidate for the redundant use case. In Fig.~\ref{fig:kappa}, we see that this choice of preconditioner performs very well, the value of the spectral condition number $\kappa$
becomes constant and independent of problem size, i.e. the number of antennas in the array. Preconditioning therefore effectively eliminates $\kappa$ from Eq.~\eqref{eq:cg_bound}. This result
is confirmed by Fig.~\ref{fig:cg_itr}. Fig.~\ref{fig:cg_itr} shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bmH$ is independent of the number of antennas in the array.
Moreover, the number of major iterations required is also much smaller than the dimension of $\bmH$.

@@

We can now use the sparsity factor to compute the order of the computational complexity of inverting $\bmH$ (assuming the effect of $\kappa$ is negligible) with
\begin{equation}
P^{c} = (1 - \gamma)P^2.
\end{equation}
Solving for $c$ we obtain
\begin{align}
c &= \log_{P}(1 - \gamma) + 2 & c_{\infty} &= \lim_{N\rightarrow \infty} c = 2. \label{eq:c}
\end{align}

@@

\section{Conjugate Gradient Method}
\label{sec:conj_grad}
The CG method was independently discovered by \citet{Hestenes1973} and \citet{Stiefel1952}. They later published a joint paper, which is now considered as the seminal
reference on CG \citep{Hestenes1952}. The iterative CG method is used to solve a particular class of linear system. The CG method is generally used to solve
\begin{equation}
\label{eq:linear_system}
\bb = \bA\bx,
\end{equation}
where $\bb\in\mathbb{C}^P$ is a known complex column vector of size $P$, $\bx\in\mathbb{C}^P$ is an unknown complex column vector with the same length as $\bb$, and $\bA\in\mathbb{C}^{P\times P}$ is a square positive-definite Hermitian matrix with the appropriate dimensions.  
Moreover, using the CG method to solve Eq.~\eqref{eq:linear_system} will only be computationally advantageous if $\bA$ is also sparse (contain many zero entries).
Using the iterative CG method to solve large sparse linear sytems was popularized by \citet{Reid1971}. A good tutorial on the CG method can be found in \citep{Shewchuk1994}.
The iterative CG method is presented in Algorithm \ref{algo:CG}.

@@

\subsection{Computational Complexity}
\citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
\citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of a major outer loop (line 6 of Algorithm \ref{algo:CG}). It turns out that at worst 
the number of major iterations that CG requires to converge\footnote{If the $\bA$-norm is used to calculate the error.} is proportional to $\sqrt{\kappa}$, where $\kappa$ is the spectral condition number of the matrix $\bA$.
The mathematical definition of $\kappa(\bA)$ is
\begin{equation}
\label{eq:kappa}
\kappa(\bA) = \frac{\lambda_{\textrm{max}}}{\lambda_{\textrm{min}}}, 
\end{equation}
where $\lambda_{\textrm{max}}$ and $\lambda_{\textrm{min}}$ respectively denote the largest and the smallest eigenvalue of $\bA$.
Moreover, the most expensive operation that is conducted within each major loop of Algorithm~\ref{algo:CG} is the vector-matrix product $\bA\bp$, which has a complexity of $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in $\bA$.
The computational complexity of CG is therefore 
\begin{equation}
\label{eq:cg_bound}
\mathbb{O}(\sqrt{\kappa}m). 
\end{equation}
It should now also be clear to the reader why the CG method is especially applicable to sparse linear systems. In sparse linear systems $m < P^2$.

\subsection{Preconditioning}
\label{sec:precon}
Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
We can now indirectly solve for $\bx$ by using 
\begin{equation}
\label{eq:prec}
\bM^{-1}\bA\bx = \bM^{-1}\bb.
\end{equation}
If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
of CG (see Eq.~\eqref{eq:cg_bound}). In this paper, we denote matrix inversion with $(*)^{-1}$. The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the identity matrix are equal to one).
If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$. However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place. 
In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).
